\documentclass{beamer}  
\usepackage{setspace}
\linespread{1.3}
\usepackage{amssymb, amsmath, amsthm}
\usepackage{appendixnumberbeamer}

% A simple slide counter at the bottom of the slide
\setbeamertemplate{footline}{
   \begin{beamercolorbox}[ht=4ex,leftskip=0.3cm,rightskip=0.3cm]{author in head/foot}
    \vspace{0.1cm}
    \hfill \insertframenumber / \inserttotalframenumber
  \end{beamercolorbox}
   \vspace*{0.1cm}
} 


%gets rid of navigation symbols
\setbeamertemplate{navigation symbols}{}

\setbeameroption{show notes}
%\setbeameroption{show only notes}
%\setbeameroption{hide notes}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[Edgeworth-98]{Improving on 97th-order Edgeworth Expansions for the Mean of a Normal Population}
\author[Hyde \& Nosferatu]{Jekyll N.\ Hyde\inst{1,3}  \and Archibald W.\ Nosferatu\inst{2}}
\institute[UTran]{\inst{1} Center for the Study of Dark Arts, University of Transylvania, \and %
                      \inst{2} Transylvania Federal Reserve Bank \and %
                      \inst{3} NBER} 

\date

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document} 
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
  \frametitle{Best Practices for Slides}
  \begin{itemize}
    \item Use a clean, simple template.
    \item Use \alert{color} effectively.
    \item Don't put too much text or math on a slide.
    \item Format graphics and tables specifically for your talk.
    \item Read the Beamer manual!
  \end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
  \frametitle{This slide is about blocks}

  \begin{block}{This is a block}
    And this is the content of the box, where you add details.
  \end{block}

  \begin{alertblock}{This is an alertblock}
    And this is the content of the alertblock, where you add details.
  \end{alertblock}


  \alert{This is an alert! It's just like regular text but more emphatic!}

\end{frame} 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{How can I remember what to say?}

  There is almost nothing on this slide, but there are hidden notes to remind me what to say! 

  \note[item]{Don't forget to mention that my favorite color is blue.}

  \note[item]{Explain Fermat's last theorem}

  \note[item]{Buy toothpaste}
\end{frame}

\end{document}

